Understanding the nanoscale structure of hexagonal phase lyotropic liquid crystal membranes

### 2018-02-01 Ben_Manuscripts/structure_paper/Draft.tex
New figures and edits

  %BJC: TODO: reproduce figure in latex (or gimp?) with higher resolution arrows
  \begin{figure}
	\centering
	\includegraphics[width=0.75\linewidth]{build.PNG} %BJC: put an xyz axis with the unit cell
	\caption{(a) We parameterized a single monomer and annealed it to produce a low energy
		configuration. (b) A python script rotates and assembles monomers into layers with 
		hydrophilic centers. (c) We chose to stack twenty layers on top of each other to create
		pores. (d) The pores are duplicated and placed into a monoclinic unit cell.}\label{fig:python}
  \end{figure}

@@

@@

  The normalized pair distribution function, $g(\mathbf{r})$, describes
  the probability of finding a pair of particles separated by $\mathbf{r}$,
  \begin{equation}
	g(\mathbf{r})= \frac{1}{\rho N} \Bigg \langle \sum_{i=1}^{N}\sum_{j\neq i}^{N} \delta(\mathbf{r}+\mathbf{r_j}-\mathbf{r_i}) \Bigg \rangle
	\label{eqn:correlation}
  \end{equation}
  where $\rho$ is the average number density of particles and $\delta(\mathbf{r})$
  is the Dirac delta function\cite{kuriabova_linear_2010}. We applied equation 
  \ref{eqn:correlation} in one and two dimensions to learn about various spatial
  correlation functions.

  In the 1D case, we measured the pair correlation function between centers of
  masses of aromatic head group rings, $g(z)$, along the z-axis (perpendicular to
  the membrane plane). To extract the average distance between layers we applied
  a discrete Fourier transform to $g(z)$ and extracted the highest intensity
  frequency.

  We looked at the xy and xz pair distribution functions based on the center of
  masses of the aromatic head group rings. We gave special consideration to the
  xy cross-sections since there are multiple rings in the same plane at different
  vertices of a polygon. We used a rotated frame of reference to avoid
  ambiguities in the final pair distribution functions. For a given center of
  mass, we rotated the coordinates so that the vector created from the center of
  mass to the pore center was in line with the positive x axis.

@@

  % BJC: I think this figure belongs somewhere else 
%MRS4: yeah, probably. Possibly supplemental information.
%  \begin{figure}
%	\centering
%	\begin{subfigure}[b]{0.32\textwidth}
%		\centering
%		\includegraphics[width=\textwidth]{sandwiched.png}
%		\caption{}\label{fig:sandwiched}
%	\end{subfigure}
%	\begin{subfigure}[b]{0.32\textwidth}
%		\centering
%		\includegraphics[width=\textwidth]{PD.png}
%		\caption{}\label{fig:pd}
%	\end{subfigure}
%	\begin{subfigure}[b]{0.32\textwidth}
%		\centering
%		\includegraphics[width=\textwidth]{Tshaped.png}
%		\caption{}\label{fig:tshaped}
%	\end{subfigure}
%	\vskip\baselineskip
%	\begin{subfigure}[b]{0.475\textwidth}
%		\centering
%		\includegraphics[width=\textwidth]{sandwichedlayers.png}
%		\caption{}\label{fig:sandwichedlayers}
%	\end{subfigure}
%	\begin{subfigure}[b]{0.475\textwidth}
%		\centering
%		\includegraphics[width=\textwidth]{offsetlayers.png}
%		\caption{}\label{fig:offsetlayers}
%	\end{subfigure}
%	\caption{(a) Sandwiched benzene dimers stack 3.8 \angstrom~apart. (b) Parallel-Displaced benzene dimers stack
%	3.4 \angstrom~vertically and 1.6 \angstrom~horizontally apart. (c) T-shaped benzene dimers stack 5.0 \angstrom~apart. 
%	(d) Two monomer layers stacked in the sandwiched configuration (e) Two monomer layers stacked in the parallel-displaced
%	configuration }\label{fig:stacking}
%  \end{figure}

  We explored the pore composition by measuring the average number densities of
  various monomer components as a function of distance from the pore centers. We
  looked at the average number density of sodium ions, aromatic rings and carbon
  atoms making up the monomer tails. We binned the radial distance of all atoms
  in each group from the pore centers, then normalized by the volume of the
  annulus defined by the bin edges and z box vector.
  %BJC: Maybe write an equation here? It's hard to say precisely how I did it
  %in words without it sounding confusing

  We calculated ionic conductivity using two different methods for robustness.
  The Nernst-Einstein relation relates the DC ionic conductivity to ion
  diffusivity, $D$, concentration, $C$, ion charge, $q$, the Boltzmann constant,
  $k_b$, and temperature, $T$: $$\sigma = \dfrac{q^2CD}{k_b T}$$ 
  We measured sodium ion diffusion coefficients by calculating the slope
  of the linear region of the z-direction mean square displacement curve as
  indicated by the Einstein relation \cite{einstein_investigations_1956}. We
  visualized the MSD plot to determine where to begin and end a linear fit. We
  measured ion concentration with respect to the volume of the entire unit cell.

  The second method, termed the Collective Diffusion model, measures the
  movement of the collective variable, Q, which is defined as the amount of
  charge transfer through the system and can be thought to represent the center
  of charge of the system. The conductance, $\gamma$, of the system can be
  calculated as: $$ \gamma = \dfrac{D_Q}{k_b T} $$ We convert the resulting value
  to ionic conductivity by multiplying by channel length and dividing by the
  membrane cross sectional area. $D_Q$ is the diffusion coefficient of the
  collective variable Q. It is calculated using the Einstein relation. One can
  access a full derivation of the model elsewhere \cite{liu_collective_2013}.

@@

  \subsection{Determining the Number of Monomers per Layer}

  Our simulations best support a model built with 5 monomers per layer based on
  the measured equilibrated pore-to-pore distances. To discern the composition of
  the monomer layers, addressing (\ref{point:monomernum}), we ran simulations of
  systems created with 4 - 8 monomers per layer. We built systems in both the
  parallel displaced and sandwiched configurations with layers initially spaced
  3.7 \AA~apart. We prepared equilibrated configurations according to the dry
  equilibration procedure. All systems are stable after 400 ns of simulation. In
  a sense, all systems are at least metastable, addressing
  (\ref{point:metastable}), however not all will make physical sense or fit the
  experimental profile that we are looking to match. Figure~\ref{fig:p2p} shows
  the pore spacing for all systems tested. Systems built with 5 monomers in each
  layer equilibrate to a pore spacing that is most consistent with the
  experimental value of 4.12 nm derived from SAXS measurements
  (Figure~\ref{fig:SAXS}). The remainder of this discussion will focus on the
  analysis of systems built with 5 monomers per layer.

  \begin{figure}
	\centering
	\includegraphics[width=\linewidth]{p2p.png}
	\caption{When we build systems with 5 monomers per layer in the
		parallel displaced configuration, the equilibrated pore spacing is closest to
		the experimental value of 4.12 nm. The equilibrated pore spacing of the model
		increases as the number of monomers in each layer increases. The pore spacing
		of systems starting in the sandwiched configuration is systematically slightly
		lower than those started in the parallel displaced configuration.}~\label{fig:p2p}
  \end{figure}

  \subsection{Layer partitioning}

  We learned that layers are well-defined and persistent, answering
  (\ref{point:layers}). We established our conclusion by plotting the one dimensional pair
  distribution function, $g(z)$, calculated between atoms along the length of the
  pores (Fig.~\ref{fig:zdf}). We measured $g(z)$ with respect to aromatic rings in
  the head groups and, separately, with respect to carbon atoms in the alkane
  chains. Using a Fourier transform of $g(z)$, we see that sandwiched
  configuration layers stack 4.39 \AA~apart while parallel displaced
  configuration layers stack 4.38 \AA~apart.

  Insert paragraph about 2D pair distribution function (yz cross section).

  We measured the correlation length, L, by fitting a function of the form:
  $$1 + Asin(Bx + C)e^{-z/L}$$ where z is the independent variable and all
  other variables are fitting parameters. The correlation length for each
  system is given in Table~\ref{fig:correlation_length}.

  \begin{table}[h]
  \centering
  \begin{tabular}{ccc}
  \toprule
  System & Layer Separation & Correlation Length \\
  \midrule
  Sandwiched & 4.39 \AA & 41 $\pm$ 3 \AA \\
  Parallel Displaced & 4.38 \AA & 86 $\pm$ 14 \AA \\
  Experiment* & 3.7 \AA & 32 $\pm$ 1 \AA \\  
  \bottomrule
  \end{tabular}
  \caption{*See supplemental information for derivation of experimental correlation length}~\label{fig:correlation_length}
  \end{table}

  %BJC: TODO: increase font size on legends
  \begin{figure}
        \centering
        \begin{subfigure}{0.45\textwidth}
                \centering
                \includegraphics[width=\textwidth]{zdf_overlay_layered.png}
                \caption{}\label{fig:zdf_layered}
        \end{subfigure}
        \begin{subfigure}{0.45\textwidth}
                \centering
                \includegraphics[width=\textwidth]{zdf_overlay_offset.png}
                \caption{}\label{fig:zdf_offset}
        \end{subfigure}
	\caption{We support the existence of layers with the observance of
		clear periodic maxima in the $z$-direction pair distribution functions of the
		(a) 5 monomer per layer, sandwiched and (b) 5 monomer per layer, parallel
		displaced configurations. The extent to which we consider a system to be
		layered is based on the magnitude of the deviation of the periodic maxima from
		the average number density. The sandwiched configuration possesses a higher
		degree of layering than the parallel displaced configuration and, in both
		cases, the head group region has a higher degree of layer partitioning than the
		tail region. (See supplemental information for detailed definitions of the head
		group and tail regions)}\label{fig:zdf}
  \end{figure}

  \subsection{Determination of Relative Interlayer Head Group Orientation}

  We answer question (\ref{point:orientation}) by simulating X-ray diffraction
  patterns produced from equilibrated MD trajectories. We tested systems built
  with 5 monomers per layer in the parallel displaced and sandwiched
  configurations. We generated simulated patterns using portions of simulation
  trajectory after equilibration. The patterns for both structures are shown and
  compared to experiment in Figure \ref{fig:XRDsim}.

@@

  \subsection{Initial Layer Spacing Affects System Equilibration}

%MRS4: can we come up with another name?  I still think disordered pores is a decent name, since there is no pi-stacking layering.  ``structures that form when layers are 5 AA apart'' is not so good.  or ``more disordered''?
%BJC3: agreed, I'll stick with disordered pores
%  The $z$-direction correlation functions show that layers in our model prefer
%  to stack further apart than 3.7 \angstrom~as suggested by experiment.
  When we build systems with layers stacked 5.0 \AA~apart and then let them
  equilibrate, we observe long-term stability of a qualitatively different
  configuration, suggesting that we have found another metastable free energy
  basin, further corroborating (\ref{point:metastable}). The basins are
  differentiated by their relative degrees of ordering resulting from different
  spatial arrangements.

  We expect that the head groups in more ordered systems will be predominantly
  coplanar with the xy plane. We defined a director vector, \textbf{n}, as the
  vector normal to the plane of the monomer's aromatic head group, then measured
  the distribution of angles between it and the z-axis. We used the resulting
  distibution of angles to calculate the nematic order parameter, S. The values
  of S for systems resulting from each initial configuration are presented in
  Table \ref{table:nematic}. The full distributions of angles between \textbf{n}
  and the z-axis are presented in the supplemental material. In general, systems
  started with layers spaced 3.7 \AA apart are more nematically ordered than
  those initially spaced 5 \AA apart. We will refer to systems built with layers
  initially spaced 3.7 \AA apart and 5.0 \AA apart as the ordered and disordered
  pore basins respectively. We studied each basin in both the parallel displaced
  and sandwiched configurations.

  \begin{table}[h]
  \centering
  \begin{tabular}{cc}
  \toprule
  System & Nematic Order Parameter (S) \\
  \midrule
  Sandwiched, Ordered & 0.659 $\pm$ 0.003 \\
  Parallel Displaced, Ordered & 0.601 $\pm$ 0.003 \\
  Sandwiched, Disordered & 0.565 $\pm$ 0.002 \\
  Parallel Displaced, Disordered & 0.503 $\pm$ 0.002 \\
  \bottomrule
  \end{tabular}
  \caption{The nematic order parameter is higher when layers are initially spaced
  3.7 \AA apart. A perfectly stacked system with aromatic rings coplanar with the
  xy plane would have a nematic order parameter of 1, while isotropically aligned
  director vectors would have a nematic order parameter of 0.}~\label{table:nematic}
  \end{table}

  We observe more evidence of disorder in the disordered pore basin when we 
  plot the two dimensional pair distribution function between the centers of 
  mass of monomer head groups. ADD MORE ANALYSIS ONCE CODE IS WORKING HOW I WANT IT

  \begin{figure}
  \begin{subfigure}{1\linewidth}
        \centering
        \begin{subfigure}{0.45\linewidth}
                \centering
                \includegraphics[width=\linewidth]{layered_xy_rings.png}
                \caption{}~\label{fig:sandwich_xy}
        \end{subfigure}%
        \begin{subfigure}{0.45\linewidth}
                \centering
                \includegraphics[width=\linewidth]{offset_xy_rings.png}
                \caption{}~\label{fig:offset_xy}
        \end{subfigure}
        \begin{subfigure}{0.45\linewidth}
                \centering
                \includegraphics[width=\linewidth]{disorder_sandwich_xy_rings.png}
                \caption{}~\label{fig:disorder_sandwich_xy}
        \end{subfigure}%
        \begin{subfigure}{0.45\linewidth}
                \centering
                \includegraphics[width=\linewidth]{disorder_offset_xy_rings.png}
                \caption{}~\label{fig:disorder_offset_xy}
        \end{subfigure}
  \end{subfigure}
  \caption{}\label{fig:xy_correlation}
  \end{figure}

  The simulated X-ray diffraction patterns indicate further structural
  differences. In the parallel displaced configuration, almost all contrast
  between R-spots and R-alkanes fades (Fig.~\ref{fig:offset_disordered_xrd}). In
  the sandwiched configuration, R-spots is weakly present, but in different
  locations, showing higher intensity at the top and bottom of the pattern as
  well as at the intersection of R-alkanes with q$_z$ = 0 (Fig.
  ~\ref{fig:layered_disordered_xrd}).

  The sandwiched and parallel displaced assemblies do not deviate from their
  initial head group arrangement. R-helix is still faintly visible in the
  parallel displaced configuration and is absent in the sandwiched simulated
  diffraction pattern. The spectroscopic signatures are unique to the two
  different head group configurations.

  The equilibrated pore spacing and distance between layers are different in
  the disordered pore basin. For both head group stacking arrangements, we
  observe a decrease in pore spacing (Fig.~\ref{fig:p2p_disordered}) and a
  corresponding increase in the equilibrated distance between layers
  (Fig.~\ref{fig:dbwl_disordered}).

@@

  \subsection{Pore Composition Depends on Initial Configuration}

@@

  In all cases, the space in the pore region is filled with sodium ions and
  head groups. In the ordered pore systems, there is less density in the center of
  the pore indicating a more ordered pore structure. In both the sandwiched and
  parallel displaced configurations, we see the density of head groups and sodium
  ions fall to less than 50\% of its maximum at r = 0
  (Fig.~\ref{fig:layered_density}). The situation is most pronounced in the
  sandwiched configuration where the maximum head group density occurs 0.44 nm
  from the pore center. The parallel displaced configuration reaches its maximum
  0.35 nm from the pore center. In contrast, both disordered pore systems show
  very little difference in density from its maximum. This implies a more uniform
  distribution of head groups within the pore center. 
  % High vacancy is less likely to be stable in the real system

@@

  In all cases, water disrupts structuring of the model \ref{fig:solvation}.
  When we add water to the system, the intensity of the reflections decrease. In
  systems built with 5 wt\% water, R-$\pi$ and R-spots become nearly
  indistinguishable from R-alkanes.

  % BJC: unsure where or if this paragraph fits
  Water is not necessary to maintain an ordered pore structure. We do not
  eliminate the possibility that water is necessary in order to drive
  self-assembly, but studying the mechanisms of self-assembly is beyond the
  scope of this work. According to our model, once the system has formed the
  Col\textsubscript{h} phase, adding water only drives disorder of the pore
  structure. In the true equilibrium configuration, if water exists, it is
  primarily confined to the pore region where there is no driving force for
  aggregation of water molecules. In the case of trace water, water molecules
  will be too sparse to form a hydrogen bonding network.

@@

  The membrane swells when we introduce water. The location of maximum head
  group density shifts from 0.35 to 0.62 nm and from 0.44 to 0.61 nm in the
  parallel displaced and sandwiched configurations respectively. Again, we
  observe the existence of ions, head groups and water outside the pore region,
  however in the hydrated system, the head groups drift beyond 1.5 \AA~from the
  pore center. In the dry systems, head groups did not wander beyond 1.4 \AA~from
  the pore center. Both observations suggest that water pushes all components
  radially outward from the pore center, characteristic of a swelling process.

>>>>

  \begin{figure}
	\centering
	\includegraphics[width=0.75\linewidth]{build.PNG} %BJC: put an xyz axis with the unit cell
	\caption{(a) We parameterized a single monomer and annealed it to produce a low energy
		configuration. (b) A python script rotates and assembles monomers into layers with 
		hydrophilic centers. (c) We chose to stack twenty layers on top of each other to create
		pores. (d) The pores are duplicated and placed into a monoclinic unit cell.}\label{fig:python}
  \end{figure}

@@

  \subsubsection{\textit{The nematic order parameter}}

  We calculated the nematic order parameter for our system in order to
  understand the degree of ordering among monomer head groups. Typically, the
  nematic order parameter is calculated for nematic liquid crystal systems which
  are characterized by unidirectional ordering of liquid crystal monomers. The
  preferred direction of monomers is defined by the unit director vector,
  $\mathbf{n}$. Assuming a single preferred direction of alignment, the nematic
  order parameter, $S$, is defined as:
  \begin{equation}
	 S = \frac{1}{2} \langle(3cos^2\theta -1)\rangle
	\label{eqn:nematic_order_parameter}
  \end{equation}
  where $\theta$ is the angle between the molecular long axis and $\mathbf{n}$.
  In a perfectly ordered system, the molecular axis should be aligned with
  $\mathbf{n}$ and give an order parameter of $S=1$. We are interested in
  quantifying the degree of monomer head group alignment between systems. We use
  Equation~\ref{eqn:nematic_order_parameter} to accomplish this by defining
  $\mathbf{n}$ as the z-axis, or pore axis, and then measuring the angles,
  $\theta$, between $\mathbf{n}$ and the vectors perpendicular to the plane of
  the aromatic head groups. See the supplemental information for a diagram.

@@

  The normalized pair distribution function, $g(\mathbf{r})$, describes
  the probability of finding a pair of particles separated by $\mathbf{r}$,
  \begin{equation}
	g(\mathbf{r})= \frac{1}{\rho N} \Bigg \langle \sum_{i=1}^{N}\sum_{j\neq i}^{N} \delta(\mathbf{r}+\mathbf{r_j}-\mathbf{r_i}) \Bigg \rangle
	\label{eqn:correlation}
  \end{equation}
  where $\rho$ is the average number density of particles and
  $\delta(\mathbf{r})$ is the Dirac delta function\cite{kuriabova_linear_2010}.
  We applied equation \ref{eqn:correlation} in three dimensions and then
  extracted one and two dimensional distribution functions by collapsing the grid
  along the appropropriate axes.

  In the 1D case, we measured the pair distribution function between centers of
  masses of aromatic head group rings, $g(z)$, along the z-axis (perpendicular to
  the membrane plane). To extract the average distance between layers we applied
  a discrete Fourier transform to $g(z)$ and extracted the highest intensity
  frequency.

  In two dimensions, we observed xy and xz pair distribution functions based on
  the center of masses of the aromatic head group rings. We gave special
  consideration to the xy cross-sections since multiple rings exist in the same
  plane located at different vertices of a polygon. We used a rotated frame of
  reference to avoid ambiguity in the final pair distribution functions. For a
  given center of mass, we rotated the coordinates so that the vector created
  from the center of mass to the pore center was in line with the positive
  x-axis.

  \subsubsection{\textit{Radial distribution functions}}
  Closely related to the spatial correlation functions, we explored the pores'
  compositions by measuring the average number densities of various monomer
  components as a function of distance from the pore centers. We looked at the
  average number density of sodium ions, aromatic rings and carbon atoms making
  up the monomer tails. We binned the radial distance of all atoms in each group
  from the pore centers, then normalized by the volume of the annulus defined by
  the bin edges and the z box vector.

@@

  \subsubsection{\textit{Ionic conductivity calculations}}

  We calculated ionic conductivity using two different methods for robustness.
  The Nernst-Einstein relation, relates the DC ionic conductivity, $\sigma$, to ion
  diffusivity, $D$, concentration, $C$, ion charge, $q$, the Boltzmann constant,
  $k_b$, and temperature, $T$: 
  \begin{equation}
	\sigma = \dfrac{q^2CD}{k_b T} 
	\label{eqn:nernst_einstein}
  \end{equation}
  We measured sodium ion diffusion coefficients by calculating the slope
  of the linear region of the z-direction mean square displacement curve as
  indicated by the Einstein relation \cite{einstein_investigations_1956}. We
  visualized the MSD plot to determine where to begin and end a linear fit. We
  measured ion concentration with respect to the volume of the entire unit cell.

  The second method, termed the Collective Diffusion model, measures the
  movement of the collective variable, Q, which is defined as the amount of
  charge transfer through the system and can be thought to represent the center
  of charge of the system. The conductance, $\gamma$, of the system can be
  calculated as:
  \begin{equation}
	 \gamma = \dfrac{D_Q}{k_b T} 
	\label{eqn:collective_diffusion}
  \end{equation}
  We convert the resulting value to ionic conductivity by multiplying by
  channel length and dividing by the membrane cross sectional area. $D_Q$ is the
  diffusion coefficient of the collective variable Q. It is calculated using the
  Einstein relation. One can access a detailed derivation of the model elsewhere
  \cite{liu_collective_2013}.

@@

  \subsection{Structural Characterization}
  \subsubsection{Number of monomer per layer}

  Our simulations best support a model built with 5 monomers per layer based on
  the measured equilibrated pore-to-pore distances. To discern the composition of
  the monomer layers, addressing Question (\ref{point:monomernum}), we ran
  simulations of systems created with 4 - 8 monomers per layer. We built systems
  in both the parallel displaced and sandwiched configurations with layers
  initially spaced 3.7 \AA~apart. We prepared equilibrated configurations
  according to the dry equilibration procedure. All systems are stable after 400
  ns of simulation. In a sense, all systems are at least metastable, addressing
  Question (\ref{point:metastable}), however not all will make physical sense or fit
  the experimental profile that we are looking to match. Figure~\ref{fig:p2p}
  shows the pore spacing for all systems tested. Systems built with 5 monomers in
  each layer equilibrate to a pore spacing that is most consistent with the
  experimental value of 4.12 nm derived from SAXS measurements
  (Figure~\ref{fig:SAXS}). The remainder of this discussion will focus on the
  analysis of systems built with 5 monomers per layer.

  \begin{figure}
	\centering
	\includegraphics[width=\linewidth]{p2p.png}
	\caption{When we build systems with 5 monomers per layer in the
		parallel displaced configuration, the equilibrated pore spacing is closest to
		the experimental value of 4.12 nm. The equilibrated pore spacing of the model
		increases as the number of monomers in each layer increases. The pore spacing
		of systems starting in the sandwiched configuration is systematically 
		lower than those started in the parallel displaced configuration.}~\label{fig:p2p}
  \end{figure}

  \subsubsection{Relative interlayer head group orientation}

  We answer Question (\ref{point:orientation}) by simulating X-ray diffraction
  patterns produced from equilibrated MD trajectories. We tested systems built
  with 5 monomers per layer in the parallel displaced and sandwiched
  configurations. We generated simulated patterns using portions of simulation
  trajectory after equilibration. The patterns for both structures are shown and
  compared to experiment in Figure \ref{fig:XRDsim}.

@@

  \subsubsection{Layer partitioning}

  We learned that layers are well-defined and persistent, answering Question
  (\ref{point:layers}). We established our conclusion by plotting the two
  dimensional pair distribution function, $g(y,z)$ of head group phenyl rings
  using Equation~\ref{eqn:correlation} (Figure \ref{fig:yz_correlation}). In both
  configurations, the phenyl rings stack in the z-direction, along $y=0$, and
  clear layers are formed. The layers in the sandwiched configuration are more
  defined than those in the parallel displaced configuration and are strongly
  correlated with adjacent layers. We also observe correlation of layers between
  pores indicated by the non uniformity in $g(y,z)$ in adjacent columns (along $y
  = \pm 3.65$).Parallel displaced layers display weaker correlations but do not
  decay in their intensity as z increases. We do not see correlation between
  pores, however correlations within pores are strong throughout the length of
  the unit cell since there is no noticeable decay in layer intensity.

  \begin{figure}
  \centering
  \begin{subfigure}{0.45\textwidth}
	\centering
	\includegraphics[width=\textwidth]{layered_yz_correlation.png}
	\caption{}\label{fig:layered_yz_correlation}
  \end{subfigure}
  \begin{subfigure}{0.45\textwidth}
        \centering
        \includegraphics[width=\textwidth]{offset_yz_correlation.png}
        \caption{}\label{fig:offset_yz_correlation}
  \end{subfigure}
  \caption{2D pair distribution functions generated from systems built in the 
	   sandwiched (a) and parallel displaced (b) configurations, provide
           a more detailed picture of the pore structure in each. The sandwiched
	   configuration}\label{fig:yz_correlation}
  \end{figure}

  We plotted the one-dimensional pair distribution function, $g(z)$, for the
  phenyl rings and for carbon atoms that make up the monomer tails (Figure
  \ref{fig:zdf}). The plots show clear periodic maxima where layers are present.
  Layering is most pronounced in the head group region, however there is still a
  degree of layering that persists into the tail region. Using a Fourier
  transform on $g(z)$ from the head groups, we see that sandwiched configuration
  layers stack 4.39 \AA~apart while parallel displaced configuration layers stack
  4.38 \AA~apart. Both stacking distances are in agreement with those observed
  in the simulated XRD patterns.

  \begin{figure}
        \centering
        \begin{subfigure}{0.45\textwidth}
                \centering
                \includegraphics[width=\textwidth]{zdf_overlay_layered.png}
                \caption{}\label{fig:zdf_layered}
        \end{subfigure}
        \begin{subfigure}{0.45\textwidth}
                \centering
                \includegraphics[width=\textwidth]{zdf_overlay_offset.png}
                \caption{}\label{fig:zdf_offset}
        \end{subfigure}
	\caption{We support the existence of layers with the observance of
		clear periodic maxima in the $z$-direction pair distribution functions of the
		(a) 5 monomer per layer, sandwiched and (b) 5 monomer per layer, parallel
		displaced configurations. The extent to which we consider a system to be
		layered is based on the magnitude of the deviation of the periodic maxima from
		the average number density. The sandwiched configuration possesses a higher
		degree of layering than the parallel displaced configuration and, in both
		cases, the head group region has a higher degree of layer partitioning than the
		tail region. (See supplemental information for detailed definitions of the head
		group and tail regions)}\label{fig:zdf}
  \end{figure}

  \subsubsection{Initial Layer Spacing Affects System Equilibration}

  All systems up to this point were built with layers stacked 3.7 \AA~apart.
  When we build systems with layers stacked 5.0 \AA~apart and then let them
  equilibrate, we observe long-term stability of a qualitatively different
  configuration, suggesting that we have found another metastable free energy
  basin, further corroborating Question (\ref{point:metastable}). 
  %The basins are
  %differentiated by their relative degrees of ordering resulting from different
  %spatial arrangements. 
  For reasons we will develop in the next segment of our
  discussion, configurations intially built with layers stacked 5.0 \AA~apart 
  will be grouped into the disordered pore basin, while those built with layers
  initially stacked 3.7 \AA~apart will be grouped into the ordered basin.

  The equilibrated pore spacing and distance between layers are different in
  the disordered pore basin. For both head group stacking arrangements, we
  observe an increase in the equilibrated distance between layers
  (Fig.~\ref{fig:dbwl_disordered}) and a corresponding decrease in pore spacing
  (Fig.~\ref{fig:p2p_disordered}).

  The simulated X-ray diffraction patterns indicate further structural
  differences. In the parallel displaced configuration, almost all contrast
  between R-spots and R-alkanes fades (Figure~\ref{fig:offset_disordered_xrd}).
  In the sandwiched configuration, R-spots is weakly present, but in different
  locations, showing higher intensity at the top and bottom of the pattern as
  well as at the intersection of R-alkanes with q$_z$ = 0
  (Figure~\ref{fig:layered_disordered_xrd}).

% BJC: in the interest of being concise, I leave this out for now.
%  The sandwiched and parallel displaced assemblies do not deviate from their
%  initial head group arrangement. R-helix is still faintly visible in the
%  parallel displaced configuration and is absent in the sandwiched simulated
%  diffraction pattern. The spectroscopic signatures are unique to the two
%  different head group configurations.

@@

  %BJC: this may belong in methods, but I think it might fit here because it
  % gives context for the choice of functional form to fit

  We quantified order in the z-direction by measuring the correlation lengths
  of all configurations. We measured the correlation length, L, by fitting $g(z)$
  using a non-linear least squares algorithm subject to a function of the form:
  \begin{equation}
  	1 + Asin(Bx + C)e^{-z/L}
        \label{eqn:correlation_fit}
  \end{equation}
  where z is the independent variable and all other variables are fitting
  parameters. The correlation length for each system is given in
  Table~\ref{table:nematic}.

  Correlation lengths are shorter in the disordered pore basin than in the
  ordered pore basin. Layers in the ordered pore basin are strongly correlated
  with neighboring layers up to more than 10 layers away. This behavior deviates
  from experimental evidence. The experimental correlation length suggests that
  layers are correlated with about 3 neighboring layers which is closer to the
  values exhibited by the disordered pore basin. We may need to simulate a system
  with many more layers in order to reduce the correlation contributed by
  periodic copies of the system.

% BJC: below is summarized in methods
%  We expect that the head groups in more ordered systems will be predominantly
%  coplanar with the xy plane. We defined a director vector, \textbf{n}, as the
%  vector normal to the plane of the monomer's aromatic head group, then measured
%  the distribution of angles between it and the z-axis. We used the resulting
%  distibution of angles to calculate the nematic order parameter, S. 
  We calculated the nematic order parameter for each system using Equation
  \ref{eqn:nematic_order_parameter}. The calculated values of S are presented in
  Table \ref{table:nematic}. The full distributions of angles between director
  \textbf{n} and vectors orthogonal to monomer phenyl rings are presented in the
  supplemental material. In general, ordered pore basin systems are more
  nematically ordered than disordered pore basin systems. The head groups of the
  disordered pore basin have more rotational freedom about the xy plane likely
  due to the larger separation between layers.

  \begin{table}[h]
  \centering
  \begin{tabular}{ccc}
  \toprule
  System & Correlation Length (\AA) & Nematic Order Parameter (S) \\
  \midrule
  Sandwiched, Ordered & 41 $\pm$ 3 & 0.659 $\pm$ 0.003 \\
  Parallel Displaced, Ordered & 86 $\pm$ 14 & 0.601 $\pm$ 0.003 \\
  Sandwiched, Disordered & 17 $\pm$ 3 & 0.565 $\pm$ 0.002 \\
  Parallel Displaced, Disordered & 9 $\pm$ 1 & 0.503 $\pm$ 0.002 \\
  Experiment & 10 $\pm$ 1 & -- \\
  \bottomrule
  \end{tabular}
  \caption{The nematic order parameter, S, is higher when layers are initially spaced
  3.7 \AA apart. In general, the sandwiched configuration has higher nematic order
  than the parallel displaced configuration. Correlation lengths are higher in
  the ordered pore basin. Correlation lengths in the disordered pore basin are
  in closer agreement with experiment.}~\label{table:nematic}
  \end{table}

%  \begin{table}[h]
%  \centering
%  \begin{tabular}{ccc}
%  \toprule
%  System & Layer Separation & Correlation Length \\
%  \midrule
%  Sandwiched & 4.39 \AA & 41 $\pm$ 3 \AA \\
%  Parallel Displaced & 4.38 \AA & 86 $\pm$ 14 \AA \\
%  Experiment* & 3.7 \AA & 10 $\pm$ 1 \AA \\  
%  \bottomrule
%  \end{tabular}
%  \caption{Uncertainties are derived from the diagonal entries in the
%	  covariance matrix calculated from a non-linear least squares fit to the data.
%	  *See supplemental information for derivation of experimental correlation
%	  length}~\label{table:correlation_length}
%  \end{table}

  We observe evidence of disorder in the xy plane of the disordered pore basin when
  we plot the two dimensional pair distribution function, $g(x,y)$, between the
  centers of mass of monomer phenyl rings (Figure \ref{fig:xy_correlation}). The
  difference is most apparent in the plots of the sandwiched configurations. In
  the ordered pore basin, head groups primarily stay stacked on top of each other
  while in the disordered pore basin, head groups exhibit more motion about the pore
  center and even drift into the middle of the pore. $g(x,y)$ for the parallel displaced
  configuration is less informative since adjacent layers are displaced in a way that
  occupies the gaps that are seen in $g(x,y)$ for the sandwiched configuration

  \begin{figure}
  \begin{subfigure}{1\linewidth}
        \centering
        \begin{subfigure}{0.47\linewidth}
                \centering
                \includegraphics[width=\linewidth]{layered_xy_correlation.png}
                \caption{Sandwiched, Ordered Pore Basin}~\label{fig:sandwich_xy}
        \end{subfigure}%
        \begin{subfigure}{0.47\linewidth}
                \centering
                \includegraphics[width=\linewidth]{offset_xy_correlation.png}
                \caption{Parallel Displaced, Ordered Pore Basin}~\label{fig:offset_xy}
        \end{subfigure}
        \begin{subfigure}{0.47\linewidth}
                \centering
                \includegraphics[width=\linewidth]{disorder_xy_correlation.png}
                \caption{Sandwiched, Disordered Pore Basin}~\label{fig:disorder_sandwich_xy_correlation}
        \end{subfigure}%
        \begin{subfigure}{0.47\linewidth}
                \centering
                \includegraphics[width=\linewidth]{disorder_offset_xy_correlation.png}
                \caption{Parallel Displaced, Disordered Pore Basin}~\label{fig:disorder_offset_xy_correlation}
        \end{subfigure}
  \end{subfigure}
  \caption{$g(x,y)$ for sandwiched and parallel displaced configurations in the
	  ordered and disordered pore basins illustrate disorder present in the xy plane
	  when initial configuration layers are stacked 5\AA
	  apart.}\label{fig:xy_correlation}
  \end{figure}

  \subsubsection{Chemical composition of pore columns}

@@

  In all cases, the space in the pore region is filled with sodium ions and
  head groups. Systems in the ordered pore basin are less dense in
  the center of the pore. In both the sandwiched and parallel displaced
  configurations, we see the density of head groups and sodium ions fall to less
  than 50\% of its maximum at r = 0 (Fig.~\ref{fig:layered_density}). The
  situation is most pronounced in the sandwiched configuration where the maximum
  head group density occurs 0.44 nm from the pore center. The parallel displaced
  configuration reaches its maximum 0.35 nm from the pore center. In contrast,
  both disordered pore systems show very little difference in density from its
  maximum. This implies a more uniform distribution of head groups within the
  pore center. 
  % High vacancy is less likely to be stable in the real system

@@

  In all cases, water disrupts structuring of the model.  When we add water to
  the system, the intensity of the reflections decrease. In systems built with 5
  wt\% water, R-$\pi$ and R-spots become nearly indistinguishable from R-alkanes.

@@

  The membrane swells when we introduce water. The location of maximum head
  group density shifts from 0.35 to 0.62 nm and from 0.44 to 0.61 nm in the
  parallel displaced and sandwiched configurations respectively. Again, we
  observe the existence of ions, head groups and water outside the pore region,
  however in the hydrated system, the head groups drift beyond 1.5 \AA~from the
  pore center. In the dry systems, head groups did not wander beyond 1.4 \AA~from
  the pore center. Both observations suggest that water pushes all components
  radially outward from the pore center, characteristic of a swelling process.  
  This system is a closer representation of the H\textsubscript{II} phase which
  is typically synthesized with ca. 8 wt\% water. Further investigation of
  hydrated systems can help unravel the mechanisms for selective transport in
  separations of aqueous solutions.
